---
title: |
  | \LARGE{Nation of Immigrants}
  | \vspace{0.1cm} \large{A Longitudinal Study of the Relationship of Immigration Laws with Changes in Demographics and GDP in the U.S.}
author: "Jose Zuniga"
date: "May 14, 2018"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  word_document:
    toc: yes
    toc_depth: 4
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
  fontfamily: tnr
  fontsize: 12pt
  geometry: margin=1in
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyhf{}
- \lhead{DATA 698.04}
- \rhead{Jose Zuniga}
- \rfoot{Page \thepage}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

# Abstract

Immigrants come to the United States looking to benefit from opportunities available in the U.S. labor market that they are unable to find in their country of origin. The current presidential administration claims that immigration reforms focusing on merit-based admissions and enforcement of existing laws will improve the economy. Political opponents claim the proposed policies have racial undertones and are meant to reduce the diversity of Immigrants. This paper conducts correlational research to describe the historic relationships between legislation on immigration, demographics, and the economy. Based on this empirical research, conclusions are then drawn on what is the likely outcome of the proposed immigration reforms.

# Introduction

## Background

Since the United States passed its first immigration law in 1790,[^1] the U.S. Congress has passed over 200 other immigration laws and amendments an average of once every year.[^2] Legislation has been exclusionary in some cases such as the Chinese Exclusion Act of 1882, and inclusionary on other occasions such as the Chinese Exclusion Repeal Act of 1943. The last immigration law passed was the Javier Vega, Jr. Memorial Act of November 2, 2017. It was a law naming a Border Patrol checkpoint in Texas. The impact of laws varies, but debates are currently centering around the immigration reforms being proposed by the new presidential administration. The White House states the policy changes will focus on enforcement, DACA legalization, sponsorship limitations, and Visa restrictions.[^3] The intent behind these reforms is a currently controversial topic, but a direction change is definitely happening. On February 23, 2018 the United States Citizenship and Immigration Services removed the phrase "nation of Immigrants" from its mission statement.

|Former Mission Statement                       |Current Mission Statement                       |
|:----------------------------------------------|:-----------------------------------------------|
|USCIS secures America's promise as a *nation of Immigrants* by providing accurate and useful information to our customers, granting immigration and citizenship benefits, promoting an awareness and understanding of citizenship, and ensuring the integrity of our immigration system.[^4]|U.S. Citizenship and Immigration Services administers the nation's lawful immigration system, safeguarding its integrity and promise by efficiently and fairly adjudicating requests for immigration benefits while protecting Americans, securing the homeland, and honoring our values.[^5]|

## Research Question

The presidential administration claims that immigration reforms will improve the economy through merit-based admissions and enforcement of existing laws. Political opponents claim the proposed policies have racial undertones and are meant to reduce the diversity of Immigrants. How have immigration reforms impacted the economy and demographics of the United States historically what is the likely outcome of the proposed immigration reforms?

## Theory and Hypothesis

In the current political climate, restrictions on immigration are politicized as means to protect the average American works from the influx of low-cost labor. It is true that Immigrants do come to the United States looking to benefit from opportunities available in the U.S. labor market that they are unable to find in their country of origin. It is also true that standard economic theories about competitive markets posit that immigration increases the supply of labor which in turn decreases the cost of labor (wages). This claim however, has not been supported by empirics. The impact of immigration on wages has been found to be either nonexistent or small; even when the influx was large and sudden (Kaufman, Hotchkiss 2006). This contradiction has led to explanations such as those claiming that immigrants are not perfect substitutes for native workers due to language. It has also led to counter studies where the data were dredged until the theorized negative impact was found. Although the impact of immigration on wages is debatable, economic theory is unambiguous about the positive relationship between immigration and economic growth (Kaufman, Hotchkiss 2006). The positive relationship that Immigrants share with various sectors of the U.S. economy holds regardless of the demographics of the immigrant. The is why the proposed immigration restrictions based on demographics are not likely to have a positive impact to the economy. From an empirical and historical perspective, the proposed restrictions on immigration appear to be motivated by desires to reduce the rate of racial and ethnic diversity in America. The results support this theory in showing a lack of meaningful correlation with changes in the legal flow of immigrants and economic growth over America’s history.

## Significance

The topic is important because of the significance of the United States' economy on the world stage. It is also important because movements opposing demographic shifts are growing in the United States and many news outlets are conflating the proposed immigration policy changes with these factions.

## Originality

There are descriptive studies on immigration laws, demographics in the United States, and the economy. Yet there are not as many that look at the relationship between immigration laws, demographics, and the economy throughout history. There are studies by the U.S. Census Bureau on historic demographics, but they do not address immigration law or the economy. The Migration Policy Institute[^6] has reports on immigration flows based on data from the U.S. Department of Homeland Security, but this information is also mostly demographic. A recent study was published in the Washington Post on the impact of the proposed immigration reforms on demographics and the economy, but it does not address anything prior to 2015. This research project looks at the relationship between immigration laws, demographics, and the economy. at a very granular level.

# Research Materials

## Prior Research

The U.S. Census Bureau has a demographic study examining the Foreign-born population by country from 1850 to 2010.[^7] Interesting revelations in this study are the steady decline in the Foreign-born population from 1910 to 1970 and then steady increase thereafter. In the last period examined, 2010, the Foreign-born population is still not as high as the period between 1860 and 1910. The study breaks down Natives and Foreign-born by State, Country or Origin, education level, preferred language, participation in the labor force, occupation, income, and age.

An article published by The U.S. National Archives in 2002 takes an in depth look at demographics and immigration law. *Race, Nationality, and Reality* goes into great detail about how immigration policy has a long history of trying to exclude Immigrants based on race. Immigration and nationality were legislated separately in the early years of the United States. Early Immigration laws focused on keeping out the poor and sick. Early Nationality laws focused on granting "free white persons" who immigrated from Europe citizenship while excluding Blacks and untaxed Native Americans. Federal citizenship was not even seen as that important until after the Civil War. So, when immigration and naturalization were placed under the administration of one agency, the coordination and uniform application of different laws designed to restric race (as defined by color) and geographic origin became a problem. Attempts to mitigate the confusion by focusing on ethnicity only exacerbated the problems. Citizenship was being granted and denied based on color, origin, or ethnicity depending on the case. Several pieces of legislation tried, but failed, to codify the "common understanding" of who was entitled to citizenship. Exceptions either for a against citizenship became commonplace. The lack of clarity and guidance was resulting in courts flooded with appeals. "There was no settled opinion on what constituted evidence of race [and] as long as nationality law contained racial requirements for naturalization and immigration law excluded those ineligible to naturalize," there was no way to stop the problem. Then, in 1952, immigration and naturalization laws were finally combined into a more cogent form with racial requirements completely removed.[^8] The article does a good job of showing the long history of the racism underlying historic immigration policy.

Another study published in the Washington Post last on February 26, 2018 looks at how the cuts to legal immigration proposed by the administration will impact demographics. The report states that the "Trump immigration plan could keep whites in U.S. majority for up to five more years...[by] greatly slashing the number of Hispanic and black African Immigrants entering America." This study looks at the proposed immigration policies in combination with population projections, fertility figures, and visa backlog information. It concludes by postulating that the proposed reforms will decrease the size of the population, which will therefore decrease the size of the workforce, which will therefore decrease the size of the economy.[^9] 

## Definitions

The U.S. Census Bureau defines Natives as natural born citizens and children born abroad to citizen parent(s). Foreign-born are defined as permanent residents, naturalized citizens, authorized migrants, and unauthorized migrants. The U.S. Department of Homeland Security refers to Foreign-born as aliens and subdivides their population into Immigrants, Nonimmigrants, and Inadmissible. Immigrants are foreigners permanently relocating to the U.S. and Nonimmigrants are foreigners relocating to the U.S. temporarily.[^10] Foreigners who are Inadmissible are those "not permitted by law to enter or remain in the United States." A visa is an authorization to enter the United States. "A citizen of a foreign country who seeks to enter the United States generally must first obtain a U.S. visa." 

## Primary Data Sources

### U.S. Census Bureau

The U.S. Census Bureau has wealth of information on their website[^11] regarding the U.S. population going back as far as 1790.[^12] The data compiled on race and Hispanic origin consists of the *Historical Census Statistics on Population Totals by Race and by Hispanic Origin*[^13] which spans 1790 to 1990 on race and 1850 to 1990 on Hispanic origin. Hispanic origin is defined as Spanish mother tongue, culture, surname, or origin regardless of race. More recent observations of these data are available in the *Profile of General Demographic Characteristics* (DP-1) data tables from the 2000 Census[^14] and the 2010 Census.[^15] The most recent observations of these data, from 2011 to 2016, are found in the 1-year *American Community Survey Demographic and Housing Estimates* (DP5) data tables.

The data on Foreign-born population comes from tables in the *Historical Census Statistics on the Foreign-born Population of the United State*s[^16] which spans 1850 to 2000. The Nativity of the Population[^17] table outlines the number of persons that are native and Foreign-born. The *Region of Birth of the Foreign-born Population*[^18] table breaks down the origin of the Foreign-born population into Europe, Asia, Africa, Oceania, Latin America, and Northern America. The details on which countries are sub-regions of these macro geographic regions is outlined on pages 46-47 of the United Nations 2002 Demographic Yearbook.[^19] The most recent observations of these data, from 2011 to 2016, are found in the 1-year *Selected Characteristics of the Native and Foreign-born Populations* (S0501) data tables.

The dataset compiled from all these Census data is quite impressive, but it is not yet ready for use. Several nuances need must be addressed. Notes from the metadata explain that: (1) Although the Census includes data on the free and slave Black population, "information on nativity was not collected for slaves." (2) "The attempt to enumerate all American Indians started in 1890" with earlier estimates being inaccurate because "American Indians not taxed (i.e., living in tribal society) were not included in the enumeration." (3) The 1940 Hispanic population figure only included the "White population of Spanish mother tongue". (4) In 1940 and 1950, there is Foreign-born data missing "because data on the Foreign-born population by country of birth in census publications for these years are limited almost entirely to the White population." Addressing the cause of these discrepancies is not a subject of this paper, but they are quite remarkable. It is not unexpected however, given the history of the hegemony in the United States.

To address the discrepancy caused by (1), the number of slaves disembarking in the United States[^20] is used as a proxy for Foreign-born Africans entering the United States from 1790 to 1880. To address (2), the American Indian population figures are backfilled using the percent change[^21] of the American Indian population starting from the earliest period without discrepancies which is 1890. Native born and total population figures are also adjusted to refelect the revised American Indian population figures. To address (3), the value was removed since it was not representative of all Hispanics. Then, to complete the dataset, imputation is used. The missing data outlined in (4) for the Foreign-born in 1940 and 1950 is filled in using linear interpolation between 1930 and 1960. 


$$y=y_{ 1 }+\left( x-x_{ 1 } \right) \left( \frac { y_{ 2 }-y_{ 1 } }{ x_{ 2 }-x_{ 1 } }  \right)$$

The remainder of the missing data is filled in using linear regression based on the natural log of the total population and the natural log of the variable being imputed.

$$\hat { y } ={ e }^{ \left( \beta _{ 0 }+\ln { \left( x_{ i } \right)  } \beta _{ 1 } \right)  }$$

The exceptions to the imputation via interpolation and regression are the statistics on Foreign-born persons from Europe and North America. Linear regression is not suited for these nonlinear data. Instead, imputation for other regions are subtracted from the total Foreign-born population and distributed according to weights based on the historical averages for the two regions. The resulting dataset is a complete dataset[^22] from 1790 to 2016 on race, Hispanic origin, and nativity.

### U.S. Department of State

Immigrant visas are issued to foreigners permanently relocating to the U.S. and Nonimmigrant visas are issued to foreigners relocating to the U.S. temporarily. Data regarding visas is available both from the U.S. Department of Homeland Security[^23] and the U.S. Department of State.[^24] The data[^25] from the U.S. Department of State shows every available Nonimmigrant visa type[^26] by country from 1997 to 2017. This dataset has an extraordinary amount of detail, but only on Nonimmigrant visas issued by consular offices which represents only a fraction of the Nonimmigrant visas. 

### U.S. Department of Labor

Certain Immigrant and Nonimmigrant employment related visas require approval from the U.S. Department of Labor. Employers "must obtain foreign labor certification from the U.S. Department of Labor, prior to filing a petition" for these workers. The process of obtaining certification from the U.S. Department of Labor involves substantiating[^27] that "there are not sufficient U.S. workers able, willing, qualified and available to accept the job opportunity in the area of intended employment and that employment of the foreign worker will not adversely affect the wages and working conditions of similarly employed U.S. workers." The types of employment related visas subject to this certification process are:

  + **E3**: Australian Free Trade Agreement principal's family.
  + **H1B**: Temporary workers in specialty occupations.
  + **H1B1**: Chile and Singapore Free Trade Agreement aliens.
  + **H2A/H2B**: Agricultural and Nonagricultural workers.
  + **D1**: Crewman on a marine vessel or aircraft.

The data[^28] maintained by the U.S. Department of Labor on these employment related visas lists all information, including status, on every application from 2008 to 2017. There also exists an external source[^29] that has harvested data dating back to 2000. Although the U.S. Department of Labor data has an extraordinary amount of detail, it only encompasses a small number of the available employment related visas.[^30]

### U.S. Department of Homeland Security

The U.S. Citizenship and Immigration Services division of the U.S. Department of Homeland Security has compiled the most complete list of immigration legislation available for 1790 to 1996. The list is not currently available from the source[^31] but it continues to exist in secondary sources such as the Immigration Daily[^32] website and the book *Immigration, Assimilation, and Border Security* by Yoku Shaw-Taylor from 2012. The list is fairly exhaustive, but it is missing about eight pieces of legislation from a smaller list[^33] maintained by the U.S. National Archives. Legislation from 1997 on can be found by querying a database[^34] maintained by the U.S. Congress. The database contains legislation since the 93rd Congress in 1973. The legislation on immigration that has become law since the 105th Congress in 1997 from this database, when combined with the information from the U.S. Citizenship and Immigration Services and U.S. National Archives, produces complete list[^35] of immigration legislation from 1790 to 2017.

The U.S. Citizenship and Immigration Services division also maintains data on immigration application form by type (Naturalization, Employment, Family, Humanitarian, Other) and approval status from 2012 to 2017. Data on I-129 applications used for employment are available since 2009. Combining all years of available data from the U.S. Citizenship and Immigration Services results in an unremarkable dataset.[^36] The most useful data[^37] on immigration relevant to this research are the statistics produced by the U.S. Department of Homeland Security in their annual yearbooks. The data on Immigrant visas form their *Persons Obtaining Lawful Permanent Resident Status by Type and Major Class Of Admission*[^38] tables spans 1986 to 2016. The data on *Nonimmigrant visas from their Nonimmigrant Admissions by Class of Admission*[^39] tables spans 1981 to 2016 with aggregate data on Immigrant visas since 1820. There are also data on *Aliens Apprehended*[^40] from 1925 to 2016, as well as on *Aliens Removed or Returned*[^41] from 1892 To 2016. Apprehensions are foreigner arrests, Removals are compulsory based on a court order, Returns are not based on court orders. This research takes the additional step of carving out the below Immigrant and Nonimmigrant employment visas for analysis:

  + **CW1/CW2**: CNMI-only transitional workers and family.
  + **E1, E2, E2C**: Treaty traders or investors and family.
  + **E3**: Australian Free Trade Agreement principal's family.
  + **H1B**: Temporary workers in specialty occupations.
  + **H1B1**: Chile and Singapore Free Trade Agreement aliens.
  + **H1C**: Registered nurses participating in the Nursing Relief for Disadvantaged Areas.
  + **H2A/H2B/H2R**: Agricultural/Nonagricultural/Returning workers.
  + **H3/H4**: Trainees and Spouses and children of H1, H2, or H3.
  + **I1**: Representatives of foreign information media and family.
  + **L1/L2**: Intracompany transferees and family.
  + **O1/O2/O3**: Workers with extraordinary ability, assistants, and family.
  + **P1/P2/P3/P4**: Internationally recognized athletes, entertainers, or artists, and family.
  + **Q1/Q2/Q3**: Workers in international cultural exchange programs and family.
  + **R1/R2**: Workers in religious occupations and family.
  + **TN/TD**: NAFTA professional workers and family.

The full join of these Immigrant and Nonimmigrant visa data are made uniform across their combined time domain by imputing missing years. Nonimmigrant data are imputed for 1982-1984, 1986-1988, and 1997 using linear interpolation between years where the data are present: [[1981, 1985], [1985, 1989], [1996, 1998]]. The missing Immigrant data from 1981 to 1985 is filled in by first imputing total visas issued using the historic average proportion of Nonimmigrant visas relative to total visas, then by subtracting Nonimmigrant visas by the derived number of total visas. The resulting dataset[^42] has complete cases on employment and total visas for both Immigrants and Nonimmigrants from 1981 to 2016. The dataset also includes Removals since 1892, Apprehensions since 1925, and Returns since 1927. No imputation was done on these three enforcement actions.

### U.S. Bureau of Economic Analysis

The U.S. Bureau of Economic Analysis has data on *Gross Output by Industry*[^43] from 1987 to 2016. The data are very granular and consists almost entirely of complete cases. The exceptions are from 1987 to 1996 in the subcategories of (1) Housing, (2) Other Real Estate, (3) General Government National Defense, (4) General Government Nondefense, and (5) Nursing and Residential Care Facilities. For the (1), (2), (3), and (4) subcategories, imputations are done using the average proportion of the child node under the parent node from the data that are present. This imputation is relatively easy since these subcategories are the only child nodes of Real Estate and General Government. The imputation for (5) is done using the average proportion of the sum of its sibling nodes which were Ambulatory Health Care Services and Hospitals. The two existing sibling nodes that the imputations are based on, are then adjusted downward to reflect the new sibling node under the Health Care parent. Failing to adjust them downward would inflate Health Care. The adjustment of the existing sibling nodes is based on the average proportion of each node relative to each other in the existing data. The data also require format changes as final step. The format they are provided in is great for display purposes, it is not amenable to analysis. The final compilation is a complete dataset[^44] from 1987 to 2016 that breaks down GDP by sector: Agriculture, Mining, Utilities, Construction, Manufacturing, Wholesale, Retail, Transportation, Information, FIRE, Service, Education, Entertainment, Other, Federal, State and Local.

# Research Methods

The dataset[^45] constructed from an arduous  data gathering and cleaning process consists of Immigration Laws and Census data from 1790 to 2016, employment visa data from 1981 to 2016, enforcement data from 1925 to 2016, and GDP by Industry from 1987 to 2016. Correlational research method is being conducted on the dataset to describe relationships between immigration law, demographics, employment visas, and the economy. The significance of the relationships between variables is evaluated using formal statistical tests.

```{r warning=F, message=F, fig.align='center', cache=TRUE}
github <- "https://raw.githubusercontent.com/jzuniga123"
file <- "/SPS/master/DATA%20698/DATA698_Project.csv"
data698 <- read.csv(paste0(github, file), header=1)
data698 <- data698[order(data698$Year), ]
data698 <- ts(data698[, -1], frequency=1, 
              start=min(data698[, 1]), 
              end=max(data698[, 1]))
```

## Data Characteristics

### Legislation

Historic immigration law numbers have an interesting pattern. The period between 1790 and 1931 saw an average of 3.03 pieces legislation per decade. The relative frequency of legislation did increase around the time of Mexican-American War in 1846, but the number of laws passed continued to remain relatively low. Then in 1932, in the midst the Great Depression, the number and frequency immigration laws started steadily increasing. Since 1932 there have been an average of 2.06 immigration laws being passed every year. Looking only at the period since 1997 shows an average of 3.65 immigration laws passed annually. The last 21 years have seen legislation on immigration being enacted at a frequency over 10 times greater than what was witnessed during the first 141 years of Federal legislative action on immigration. 

```{r warning=F, message=F, fig.align='center', cache=TRUE}
Laws <- data698[, 1]
mean(window(Laws, end = 1931))
mean(window(Laws, start = 1932))
mean(window(Laws, start = 1997))
```

Notable periods of increased legislative activity that were outliers in their era are:

  + **1798**: Three laws passed in response to the Quasi-War with France.
  + **1882**: Two laws passed, one being the Chinese Exclusion Act.
  + **1887**: Two laws passed restricting labor contracts and property rights.
  + **1903**: Two laws passed, one restricting anarchists.
  + **1932**: Four laws passed, three restricting labor contracts.
  + **1940**: Four laws passed, three enhancing enforcement activities.
  + **1950**: Four laws passed, three providing relief to orphans, military, and sheepherders.
  + **1978**: Seven laws passed, two restrictive to agriculture and travel, five progressive: addressing bias in admissions, expulsion of Nazis, human smuggling, and adoption.
  + **2000**: Thirteen laws passed during an election year, eleven prior to elections.
  + **2002**: Ten laws passed in the wake of the September 11, 2001 terrorists attacks.
  + **2008**: Six laws passed during an election year, all prior to elections.

```{r warning=F, message=F, fig.align='center', cache=TRUE}
library(fma)
mov_avg <- ma(Laws, order=10)
plot(Laws)
lines(mov_avg, col=2)
abline(h=c(0.75, 1.5, 2.5), col=c(3:5))
```

### Population

#### Race and Hispanic Origin (RAHO)

The purely numerical data from the decennial Census on race and Hispanic origin has two interesting features. As is usual with most social phenomena, the distributions appear mostly exponential. The two exceptions are the White race which can be described as linear since around 1870 and the Native American race which has a positive parabolic shape with a minimum at the year 1900. Note that for aggregate analysis purposes, the free and enslaved Black population that the Census counts separately until 1860, is being counted together. When the Census data on race and Hispanic origin are viewed as a percentage of total population however, some striking details emerge. Hispanics, Asians, Pacific Islanders, and Other races grow exponentially as a percentage of total population over time. Native Americans have decreased exponentially as a percentage of total population over time, from 14.2% in 1790 to 0.7% in 2016. The Black percentage of the population has decreased overall but in a cubic manner with a minimum of 9.5% in 1930 and two local maximas of 17.7% in 1810 and 11% in 1990. The White percentage of the population has a negative parabolic shape with a maximum of 89% in 1920 and an all time low of 61.7% in 2016.

```{r warning=F, message=F, fig.align='center', cache=TRUE}
library(zoo)
Census <- data698[, 2:17]
RAHO <- Census[, 2:8]
RAHO[, 2] <- RAHO[, 2] + RAHO[, 3]
RAHO <- RAHO[, -3]
Population <- rowSums(RAHO)
Proportions <- RAHO / Population
plot(na.approx(RAHO), type="l", main="Race and Hispanic Origin")
plot(na.approx(Proportions), type="l", main="Race and Hispanic Origin")
na.omit(data.frame(Year=as.numeric(time(Proportions)), round(Proportions, 6)))
```

#### Native and Foreign-born (NAFB)

The decennial numbers on the Native and Foreign-born population also follow what can be described as an exponential pattern. There is a decline in the Foreign-born population from 1940 to 1970, but by 1980 the number returns to its prior levels and continues on its prior trajectory. The proportion of Native and Foreign-born persons in the population are mirror images of each other. The median proportion of Foreign-born persons in the United States since 1790 has been 12.4%. In 1850 and 1970 the percentage of Foreign-born persons in the United States reaches lows of 9.5% and 4.8%, respectively. The proportion of Foreign-born persons in the United States has peaks in 1870, 1890, and 1910 at 14.3%, 14.8%, and 14.7%; respectively. In 2016 the percentage of Foreign-born sits at 13.5%. 

```{r warning=F, message=F, fig.align='center', cache=TRUE}
NAFB <- Census[, 9:10]
Population <- rowSums(NAFB)
Proportions <- NAFB / Population
plot(na.approx(NAFB), type="l", main="Native and Foreign-born")
plot(na.approx(Proportions), type="l", main="Native and Foreign-born")
median(Proportions[,2], na.rm=T)
na.omit(data.frame(Year=as.numeric(time(Proportions)), round(Proportions, 6)))
```

#### Foreign-born Country of Origin (FBCO)

The decennial numbers on country of origin for Foreign-born are a little more interesting. Persons from Africa, Asia, Oceania, and Latin America arrive at an exponential rate. Arrivals from Europe and North America however, show exponential growth until around 1930, and then steady declining thereafter. The proportion of Foreign-born persons in the United States by country of origin shows exponential increases from Asia, Oceania, and Latin America just like the pure numbers. The large increases start around 1860 for Oceania, and 1970 for Latin America and Asia. Foreign-born persons from Africa follow a less natural pattern due the slavery. The percentage of Foreign-born persons from Africa was high from 1790 to 1820, but relatively low thereafter. Looking at Foreign-born persons from Africa in the post-slavery era shows a more natural exponential pattern with a large increase starting in 1980. The proportion of Foreign-born persons from Europe remained above 80% until 1950 when a permanent and rapid decline started leading to its current low of 10.9% in 2016. The proportion of Foreign-born persons from North America has a very strange and unnatural pattern hovering around 8.9% until 1970 when a rapid decline started leading to its current low of 1.8% in 2016. 

```{r warning=F, message=F, fig.align='center', cache=TRUE}
FBCO <- Census[, 11:16]
Population <- rowSums(FBCO)
Proportions <- FBCO / Population
plot(na.approx(FBCO), type="l", main="Native and Foreign-born")
plot(na.approx(Proportions), type="l", main="Native and Foreign-born")
median(Proportions[,6], na.rm=T)
na.omit(data.frame(Year=as.numeric(time(Proportions)), round(Proportions, 6)))
```

### Immigration

#### Visas

The numeric data on Immigrant Visas since 1820 is best described as chaotic. It has wild swings up and down starting with an upward trend from 1820 to about 1910, then a drastic decline until about 1945, then another trend upward with a similar trajectory as before until 2016. The graph also shows a huge spike around 1990. Employment related Immigrant visas over the last 30 years are much simpler. The number of Immigrant visas related to employment held steady from 1981 to 1991, was chaotic from 1992 to 2007, and then steady ever since. There were large decreases around 1999 and 2004, and large increases around 1993, 2001, and 2006. Total immigrant visas follow a similar pattern with the exception of the huge spike around 1990 mentioned earlier. Nonimmigrant visas have a less interesting pattern. Employment related visas have a steadily increasing trend with only a slight amount of variability. All Nonimmigrant visas also have an increasing trend, but with less variability. Nonimmigrant visas make up between 89% to 95% off all visas for the last 30 years. 

```{r warning=F, message=F, fig.align='center', cache=TRUE}
plot(data698[, 19], ylab="Imm_All", main="Immigrant Visas")
Visas <- data698[, 18:21]
plot(window(Visas, start = 1981))
```

#### Enforcement

Enforcement data has some interesting patterns. Removals remained at a constant rate until 1996 when they increased substantially. Apprehensions and Returns follow almost identical patterns. There was a large increase in enforcement in the ten-year period between 1945 and 1955 with its peak being reached in 1954. The increase started again in 1965 and has remained elevated since with peaks in 1986, 1996, and 2000. Around the time Removals started steadily increasing, Apprehensions and Returns started steadily decreasing. The three enforcement actions combined follow the pattern of the latter two enforcement actions which make up the bulk of the actions. 

```{r warning=F, message=F, fig.align='center', cache=TRUE}
Enforce <- data698[, 23:25]
plot(window(Enforce, start = 1925))
```

### GDP by Industry

GDP by industry over the past 30 years has increased nearly uniformly across industries. The exceptions are Agriculture Mining, Utilities and Construction which appear to be more susceptible to business cycles. When examining the industries in aggregate as a proportion of the economy, their shares appear to remain relatively constant over the last 30 years. The Manufacturing, FIRE (Finance, Insurance, Real Estate), Service, and Education sectors make up the bulk of GDP; in that order. Over time there has been a slight decrease in the proportion of the economy captured by the Manufacturing sector and slight increases in the FIRE and Service sectors

```{r warning=F, message=F, fig.align='center', cache=TRUE}
Economy <- data698[, 27:42]
plot(window(Economy[,1:8], start = 1987))
plot(window(Economy[,9:16], start = 1987))
GDP <- rowSums(Economy)
Proportions <- Economy / GDP
plot(window(Proportions[,1:8], start = 1987))
plot(window(Proportions[,9:16], start = 1987))
```

## Data Relationships

Correlation measures the relationship between two variables. The most commonly used measure of correlation is Pearson[^46] correlation ($r$). Pearson correlation measures the relationship between normal linear homoskedastic variables. Measuring relationships between variables that are not normal, linear, or homoskedastic (inherently or through transformation) with the Pearson correlation formula produces misleading results. Spearman[^47] correlation ($\rho$) and Kendall[^48] correlation ($\tau$) are non-parametric measures of correlation based on monotonic rank. Spearman correlation is more computationally efficient than Kendall correlation, but less robust. For this analysis, the accuracy of Kendall correlation is being prioritized over the speed of Spearman correlation. 

$$r = \frac{\sum{(x-m_x)(y-m_y)}}{\sqrt{\sum{(x-m_x)^2}\sum{(y-m_y)^2}}} \tag{Pearson}$$

$$\rho = \frac{\sum(x' - m_{x'})(y'_i - m_{y'})}{\sqrt{\sum(x' - m_{x'})^2 \sum(y' - m_{y'})^2}} \tag{Spearman}$$

$$\tau = \frac{n_c - n_d}{\frac{1}{2}n(n-1)} \tag{Kendall}$$

## Other Considerations

### Policy Lag

There is a lag between the enactment of a law and its full implementation. Changes in laws require bureaucratic changes which are slow by design. Immigration applications are processed on an ongoing basis. Changes to immigration laws must be managed around applications that are already in progress. When looking at the relationship between legislation and other variables, it is therefore important to also look at the lagged relationships.

### Imputation

The Census data obtained is decennial from 1790 to 2010 and then annual until 2016. All other non-Census data is annual and complete from 1987 until 2016. To conduct a proper analysis of the past 30 years, imputation is required on the Census data for missing periods. This is done with linear interpolation[^49] between periods that there are data points present. Linear interpolation makes no assumption on distribution. It draws a straight line between two known points and returns a value along that line for a given point in the domain.

$$y=y_{ 1 }+\left( x-x_{ 1 } \right) \left( \frac { y_{ 2 }-y_{ 1 } }{ x_{ 2 }-x_{ 1 } }  \right)$$

# Results

## Legislative Correlations

### Social Factors

First the Kendall correlation between new, one-year old, two-year old, and three-year old immigration legislation and social factors is examined. When it comes to Race and Hispanic Origin, legislation has correlations between $\tau=0.51$ and $\tau=0.56$ for Hispanic origin and nearly every race. The exception is Native Americans which has a correlation that hoovers around $\tau=0.26$. For the Native and Foreign-born population, the correlations are also moderate. Native have a correlation that hoovers around $\tau=0.52$. Foreign-born correlation that hoovers around $\tau=0.44$. The correlations on Foreign-born Country of Origin are higher for Asia, Oceania, and Latin America which have correlations between $\tau=0.51$ and $\tau=0.53$. The correlation for Africa is around $\tau=0.4$. The correlations for Europe and North America are between $\tau=0.19$ and $\tau=0.26$ with the lower correlations figures being from Europe. All of these correlations are significant at $\alpha=0.05$.

```{r warning=F, message=F, fig.align='center', cache=TRUE}
library(psych)
Legislation <- window(cbind(Law_New=Laws, Law_1yr=lag(Laws, k=-1),
  Law_2yr=lag(Laws, k=-2), Law_3yr=lag(Laws, k=-3)), end=2016)
corr.test(na.approx(RAHO), Legislation, use="complete.obs", method="kendall")
corr.test(na.approx(NAFB), Legislation, use="complete.obs", method="kendall")
corr.test(na.approx(FBCO), Legislation, use="complete.obs", method="kendall")
```

### Economic Factors

The Kendall correlation between new, one-year old, two-year old, and three-year old immigration legislation and economic factors is examined next. The largest correlations with immigration legislation and visas can be found after two years. Although these correlations are larger, they are not very high. The correlations are between $\tau=0.21$ and $\tau=0.31$. Enforcement actions have low correlations that are fairly constant across lags. They range between $\tau=0.19$ and $\tau=0.24$. The largest correlations with immigration legislation and GDP by Industry can be found after three years. Just like before, although these correlations are larger, they are not very high. Excluding the highest and lowest correlations, these correlations range between $\tau=0.29$ and $\tau=0.33$. The highest correlation is with Utilities at $\tau=0.39$ and the lowest is with Agriculture at $\tau=0.25$. None of these correlations are significant at $\alpha=0.05$.

```{r warning=F, message=F, fig.align='center', cache=TRUE}
corr.test(Visas, Legislation, use="complete.obs", method="kendall")
corr.test(Enforce, Legislation, use="complete.obs", method="kendall")
corr.test(Economy, Legislation, use="complete.obs", method="kendall")
```

## Economic Correlations

As previously mentioned, Kendall correlations between legislation and the economy are low and statistically insignificant. This is not true for many other factors. Correlations between Race and Hispanic Origin and every industry in the Economy are high and statistically significant at $\alpha=0.05$ for every single race and Hispanics. The same is true for the Native and Foreign-born population and every industry; high and statistically significant at $\alpha=0.05$. Country of origin does not show the same level of equality however. At an $\alpha=0.05$ Asia, Africa, Oceania, and Latin America have high and statistically significant correlations with every industry, but Europe and North America have statistically insignificant correlations with every industry. There is also a dichotomy between Immigrant and Nonimmigrant visas. At an $\alpha=0.05$, Nonimmigrant visas have statistically significant correlations with every industry, but Immigrant visas have statistically insignificant correlations with every industry. Enforcement actions also have mixed results. Apprehensions and Returns have statistically insignificant correlations with every industry, but Removals have statistically significant correlations with every industry at $\alpha=0.05$. As stated earlier, Removals are compulsory actions based on a court order.

```{r warning=F, message=F, fig.align='center', cache=TRUE}
corr.test(Economy, na.approx(RAHO), use="complete.obs", method="kendall")
corr.test(Economy, na.approx(NAFB), use="complete.obs", method="kendall")
corr.test(Economy, na.approx(FBCO), use="complete.obs", method="kendall")
corr.test(Economy, Visas, use="complete.obs", method="kendall")
corr.test(Economy, Enforce, use="complete.obs", method="kendall")
```

# Discussion

The correlations between legislation on immigration and social phenomena are greater than the correlations between legislation on immigration and economic phenomena. The finding that the correlations between immigration laws and Native Americans makes sense intuitively. Native Americans are not generally impacted by legislation on immigration. The finding that lagging has little effect on correlations with social phenomena is indicative of the existence of closer temporal relationship between both. The effect of lagging on economic phenomena would indicate that a temporal element to is also part of the relationship if a relationship did exist. It was found however, that social and not economic phenomena have statistically significant correlations with legislation on immigration.

The correlations with industries in the economy contradicts several stereotypes. The finding that every race, Hispanic, Natives, and Foreign-born have a strong and statistically significant positive correlation with every industry in the economy is refreshing. It was a little interesting to see that the European and North American subsections of the Foreign-born did not have statistically significant correlations with any industry. This is likely related to European and North American arrivals having an unnatural distribution that was inflated until race stopped being a factor in admissions around 1950. This period coincides with the permanent and rapid decline by Europeans in 1950 and is temporally proximate to the permanent and rapid decline by North Americans starting 1970. It was also interesting to see how the temporary Nonimmigrant visas had a statistically significant correlation with industry while the permanent Immigrant visas had statistically insignificant correlations. Last, it was strange to see that court ordered Removals, but not other enforcement actions, had a statistically significant positive correlation with industry.

# Conclusion

Historically, legislation on immigration has had a statistically insignificant correlation with the economy. The statistically significant correlations that immigration laws legislation have had are with demographics. Although intent cannot be inferred, this research shows that it is statistically more likely that opponents of the administration are correct in their predicted outcomes. The proposed immigration reforms may affect demographics, but not the economy.

[^1]: https://www.migrationpolicy.org/research/timeline-1790
[^2]: https://www.archives.gov/files/research/naturalization/420-major-immigration-laws.pdf
[^3]: https://www.whitehouse.gov/briefings-statements/white-house-framework-immigration-reform-border-security/
[^4]: https://web.archive.org/web/20180221163615/https://www.uscis.gov/aboutus
[^5]: https://www.uscis.gov/aboutus
[^6]: https://www.migrationpolicy.org/ 
[^7]: https://www.census.gov/newsroom/pdf/cspan_fb_slides.pdf
[^8]: https://www.archives.gov/publications/prologue/2002/summer/immigration-law-1.html
[^9]: [https://www.washingtonpost.com/news/wonk/wp/2018/02/06/trump-immigration-plan](https://www.washingtonpost.com/news/wonk/wp/2018/02/06/trump-immigration-plan-could-keep-whites-in-u-s-majority-for-up-to-five-more-years/?utm_term=.2de44787d217)
[^10]: [https://help.cbp.gov/app/answers/detail/a_id/72](https://help.cbp.gov/app/answers/detail/a_id/72/~/what-is-the-difference-between-an-immigrant-visa-vs.-nonimmigrant-visa-%3F)
[^11]: https://factfinder.census.gov/faces/nav/jsf/pages/searchresults.xhtml
[^12]: https://www.census.gov/prod/www/decennial.html
[^13]: https://www.census.gov/population/www/documentation/twps0076/twps0076.html
[^14]: https://www.census.gov/prod/cen2000/dp1/2kh00.pdf 
[^15]: https://www.census.gov/history/pdf/orangenj-92017.pdf 
[^16]: https://www.census.gov/population/www/documentation/twps0081/twps0081.html 
[^17]: https://www.census.gov/population/www/documentation/twps0029/tab01.html 
[^18]: https://www.census.gov/population/www/documentation/twps0029/tab02.html 
[^19]: http://unstats.un.org/unsd/demographic/products/dyb/dyb2.htm 
[^20]: http://www.slavevoyages.org/assessment/estimates 
[^21]: https://nativestudy.wordpress.com/ 
[^22]: https://github.com/jzuniga123/SPS/blob/master/DATA%20698/US_Census_Data.csv
[^23]: https://www.uscis.gov/tools/reports-studies/immigration-forms-data 
[^24]: https://travel.state.gov/content/travel/en/legal/visa-law0/visa-statistics.html 
[^25]: https://travel.state.gov/content/travel/en/legal/visa-law0/visa-statistics/nonimmigrant-visa-statistics.html
[^26]: http://www.ustraveldocs.com/in/in-niv-typeall.asp#
[^27]: https://www.foreignlaborcert.doleta.gov/perm.cfm 
[^28]: https://www.foreignlaborcert.doleta.gov/performancedata.cfm 
[^29]: http://www.flcdatacenter.com/
[^30]: https://travel.state.gov/content/travel/en/us-visas/visa-information-resources/all-visa-categories.html 
[^31]: https://guides.law.fsu.edu/c.php?g=84898&p=546805 
[^32]: https://ilw.com/resources/ 
[^33]: https://www.archives.gov/files/research/naturalization/420-major-immigration-laws.pdf 
[^34]: https://www.congress.gov/search 
[^35]: https://github.com/jzuniga123/SPS/blob/master/DATA%20698/ImmigrationLaws_1790to2012.csv 
[^36]: https://github.com/jzuniga123/SPS/blob/master/DATA%20698/USCIS_2009to2017.csv 
[^37]: https://www.dhs.gov/immigration-statistics/yearbook
[^38]: https://www.dhs.gov/immigration-statistics/yearbook/2016/table6 
[^39]: https://www.dhs.gov/immigration-statistics/yearbook/2016/table25 
[^40]: https://www.dhs.gov/immigration-statistics/yearbook/2016/table33 
[^41]: https://www.dhs.gov/immigration-statistics/yearbook/2016/table39 
[^42]: https://github.com/jzuniga123/SPS/blob/master/DATA%20698/DHS_1820to2016.csv 
[^43]: https://www.bea.gov/iTable/iTable.cfm?ReqID=51&step=1#reqid=51&step=2&isuri=1 
[^44]: https://github.com/jzuniga123/SPS/blob/master/DATA%20698/GDPbyIndustry_1987to2016.csv  
[^45]: https://github.com/jzuniga123/SPS/blob/master/DATA%20698/DATA698_Project.csv 
[^46]: http://mathworld.wolfram.com/CorrelationCoefficient.html 
[^47]: http://mathworld.wolfram.com/SpearmanRankCorrelationCoefficient.html 
[^48]: https://onlinecourses.science.psu.edu/stat509/node/158 
[^49]: http://mathworld.wolfram.com/Interpolation.html 
