---
title: "BLS Public Data API"
author: "Jose Zuniga"
output:
  html_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Presentation Instructions

Make a five-minute presentation on any chosen topic, preferably any topic from the current week's chapter reading of *Data Science for Business*, or another topic that would be of interest to *most*. Do not just summarize the topic, go a little further, such as:

  + Discussing one or more business use cases, or
  + Showing a short example with R code (and perhaps relevant R package(s)), or
  + Providing a curated "learning path" of on-line resources to build further expertise in that topic.

You may also record your presentation instead of presenting in our meetup. Effective data scientists need to be effective presenters, so making a presentation in front of a group is strongly encouraged but not required. In DATA 607, our primary focus is on writing R code related to getting and shaping data in preparation for downstream modeling and presentations. 

***

# BLS Public Data API

The BLS Public Data Application Programming Interface (API) is an application designed to allow third party programmers, developers, and organizations to retrieve published historical time series data in `JSON` data-interchange format. Using Public Data API signatures, users can consume and manipulate raw data from all of the Bureau's surveys to create a wide range of applications that conform to W3C standards and accepted practices. The BLS Public Data API Version 1.0 does not require registration and is open for public use. The BLS Public Data API Version 2.0 does require [registration](http://data.bls.gov/registrationEngine/). API Version 2.0 offers greater query limits and allows users to request net changes, percent changes, and series description information. The API has a few known limitations specified on the website. The one which becomes most readily apparent as soon as one starts working with the API is that **users must have knowledge of BLS [Series IDs](http://www.bls.gov/help/hlpforma.htm) in order to successfully complete a request**.

|Service	                                |Version 2.0  |Version 1.0  |
|-----------------------------------------|:-----------:|:-----------:|
|Daily query limit	                      |500	        |25           |
|Series per query limit	                  |50	          |25           |
|Years per query limit	                  |20	          |10           |
|Net/Percent Changes	                    |Yes	        |No           |
|Optional annual averages	                |Yes	        |No           |
|Series description information (catalog) |Yes	        |No           |

# The `blsAPI` Package

The [`blsAPI`](https://cran.r-project.org/web/packages/blsAPI/blsAPI.pdf) Package allows users to request data for one or multiple series through the U.S. Bureau of Labor Statistics API.

```{r warning=FALSE, message=FALSE}
library(blsAPI)
library(dplyr)
library(tidyr)
```

## The `blsAPI` Function

`blsAPI(payload = NA, api.version = 1, return.data.frame = FALSE)`

The `blsAPI` function allows users to request data for one or multiple series through the U.S. Bureau of Labor Statistics API. Users provide parameters and the function returns a JSON string or data frame. The `Payload` parameter consists of a string or a list containing the series name(s) to be sent to the API. Most Series ID formats can be found [here](http://www.bls.gov/help/hlpforma.htm). To retrieve optional parameters, users must include their assigned registration key. Optional parameters are `startyear`, `endyear`, `calculations`, `annualaverage`, and `catalog` data.

```{r echo=FALSE, results="hide"}
API_Key <- "564ca020f9d4427c8d6af35f49551a1d"
```
```{r cache=TRUE}
payload <- 'LAUCN040010000000005'
single <- blsAPI(payload, api.version = 1, return.data.frame = T)
head(single)
payload <- list('seriesid'=c('LAUCN040010000000005', 'LAUCN040010000000006'))
multiple <- blsAPI(payload, api.version = 1, return.data.frame = T)
head(multiple)
payload <- list(
'seriesid' = c('LAUCN040010000000005', 'LAUCN040010000000006'),
'startyear' = 2010,
'endyear' = 2012,
'calculations' = TRUE,
'annualaverage' = TRUE,
'catalog' = TRUE,
'registrationKey' = API_Key)
parameters <- blsAPI(payload, api.version = 2, return.data.frame = T)
head(parameters)
```

## The `blsQCEW` Function

`blsQCEW(method, year = NA, quarter = NA, area = NA, industry = NA, size = NA)`

The `blsQCEW` function allows users to request quarterly census of employment and wages (QCEW) data from the U.S.
Bureau of Labor Statistics open access. Users provide parameters and the function returns a data
frame. This function is based off of the sample code developed by the BLS. The `method` parameter consists of a case insensitive string describing which type of data you want requested. Valid options are [`area`](http://data.bls.gov/cew/doc/titles/area/area_titles.htm), [`industry`](http://data.bls.gov/cew/doc/titles/industry/industry_titles.htm), and [`size`](http://data.bls.gov/cew/doc/titles/size/size_titles.htm).

```{r cache=TRUE}
area <- blsQCEW('area', year='2013', quarter='1', area='36005')
names(area)
industry <- blsQCEW('industry', year='2013', quarter='1', industry='21222')
names(industry)
size <- blsQCEW('size', year='2013', size='6')
names(size)
```

# Current BLS NAICS Basis API Products

For over 60 years, the Standard Industrial Classification (SIC) system served as the structure for the collection, presentation, and analysis of the U.S. economy. Over the years, there were numerous revisions to the SIC system. Despite these revisions, the system received increasing criticism. Developments in information services, new forms of health care provision, expansion of services, and high-tech manufacturing are examples of industrial changes that could not be studied under the SIC system. NAICS was introduced in 1997 and is periodically revised to reflect changes in the industrial structure of the U.S. and North American economy. NAICS provides a tool to ensure that economic statistics reflect the changing economy.

```{r}
BLS_Products <- read.csv(paste0("https://raw.githubusercontent.com/jzuniga123/SPS/master/",
                                "DATA%20607/BLS_Products.csv"), stringsAsFactors = F)
BLS_Products <- BLS_Products %>% filter(NAICS_SIC == "N" & HISTORIC !=1)
BLS_Products %>% select(SERIES) %>% arrange (SERIES)
examine <- function(Program) {
  product <- BLS_Products %>% filter(PROGRAM == Program)
  cat(noquote(paste0(rep("#", 75), collapse = "")),"\n")
  for (i in 1:nrow(product)) {
    payload <- list('seriesid' = product[i, "ID_EXAMPLE"], 'registrationKey' = API_Key)
    data <- blsAPI(payload, api.version = 2, return.data.frame = T)
    result <- list(PRODUCT_NAME = noquote(product[i, "SERIES"]),
                   LINKS = noquote(cbind(
                       information = c("PRODUCT_OVERVIEW", 
                                       "SERIES_DETAILS",
                                       "POPULAR_QUERIES"), 
                       link = c(product[i, "OVERVIEW"], 
                                product[i, "DEFINITIONS"], 
                                product[i, "POPULAR"]))),
                   PREVIEW = head(data))
    print(result)
    cat(noquote(paste0(rep("#", 75), collapse = "")),"\n")
  }
}
```

## Office of Compensation and Working Conditions

### Injuries, Illnesses, & Fatalities
```{r}
examine("IIF")
```

### National Compensation Survey
```{r}
examine("NCS")
```

## Office of Employment and Unemployment Statistics

### American Time Use Survey
```{r}
examine("TUS")
```

### Current Employment Statistics (National)
```{r}
examine("CES")
```

### Current Employment Statistics (State)
```{r}
examine("SEA")
```

### Current Population Survey
```{r}
examine("CPS")
```

### Geographic Profile of Employment & Unemployment
```{r}
examine("GPS")
```

### Job Openings and Labor Turnover Survey
```{r}
examine("JOLTS")
```

### Local Area Unemployment Statistics
```{r}
examine("LAU")
```

### Occupational Employment Statistics 
```{r}
examine("OES")
```

### Quarterly Census of Employment and Wages
```{r}
examine("CEW")
```

## Office of Prices and Living Conditions

### Consumer Expenditure Survey
```{r}
examine("CEX")
```

### Consumer Price Index
```{r}
examine("CPI")
```

### International Price Program
```{r}
examine("IPP")
```

### Producer Price Index
```{r}
examine("PPI")
```

## Office of Productivity and Technology

### Productivity & Costs
```{r}
examine("LPC")
```

# References

http://www.bls.gov/data/

http://www.bls.gov/ggs/#data

http://www.bls.gov/bls/naics.htm

http://data.bls.gov/cgi-bin/srgate

http://www.bls.gov/bls/senior_staff/

http://www.bls.gov/bls/api_features.htm

http://www.bls.gov/developers/api_r.htm

http://www.bls.gov/bls/proghome_a-z.htm

http://www.bls.gov/opub/mpbls/mpbls919.pdf

http://www.bls.gov/developers/api_signature.htm

http://www.bls.gov/developers/api_signature_v2.htm