---
title: "Assignment - Web APIs"
author: "Jose Zuniga"
output:
  html_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Instructions

The New York Times web site provides a rich set of APIs, as described here: http://developer.nytimes.com/docs

You'll need to start by signing up for an API key. Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.

# Article Search API

With the New York Times [Article Search API](http://developer.nytimes.com/article_search_v2.json), you can search New York Times articles from September 18, 1851 to today, retrieving headlines, abstracts, lead paragraphs, links to associated multimedia, and other article metadata. Note: In URI examples and field names, italics indicate placeholders for variables or values. Brackets [ ] indicate optional items. Parentheses ( ) are not a convention-- when URIs include parentheses, interpret them literally. 

**Note:** The value of *page* corresponds to a set of $10$ results. For example, *page* $=0$ correspondents to records $0$ to $9$; and *page* $=1$ correspondents to records $10$ to $19$. Also, ach API is limited to $1,000$ calls per day, and $5$ calls per second.
```{r echo=FALSE}
Hidden_API_Key <- "XXXXXXXXXXXX"
```
```{r cache=TRUE}
library(httr)
library(jsonlite)

query <- "CUNY School of Professional Studies"
results <- 500

NYT_Articles <- function(query, results) {
  NYT <- data <- {}
  set <- 0:(results / 10 - 1)
  for (i in 1:max(set)) {
    url <- paste0("https://api.nytimes.com/svc/search/v2/articlesearch.json?q=",
                  gsub(" ", "\\+", query), "&page=", i,"&api-key=", Hidden_API_Key)
    json <- GET(url)
    json <- rawToChar(json$content)
    json <- fromJSON(json, simplifyVector = TRUE)
    if (length(json$response[[2]]) ==0 ) { break }
    data <- data.frame(json)
    data[,12] <- data[,12][1]
    data <- data[,c(4:8,12,14,16,17,20,22)]
    NYT <- data.frame(rbind(NYT, data))
    if (i %% 5) { Sys.sleep(2) }
  }
  colnames(NYT) <- gsub("\\.1", "", colnames(NYT))
  colnames(NYT) <- gsub("\\w+\\.", "", colnames(NYT))
  return(NYT)
}
NYT <- NYT_Articles(query, results)
t(head(NYT, 1))
```

# References

http://developer.nytimes.com/article_search_v2.json#/README

http://web.stanford.edu/~cengel/cgi-bin/anthrospace/scraping-new-york-times-articles-with-r [(cached)](http://webcache.googleusercontent.com/search?q=cache:-Kr1l1uZr34J:web.stanford.edu/~cengel/cgi-bin/anthrospace/scraping-new-york-times-articles-with-r+&cd=2&hl=en&ct=clnk&gl=us)

http://web.stanford.edu/~cengel/cgi-bin/anthrospace/wp-content/uploads/2009/09/scrapeNYT_API2.txt