---
title: "Project - Generate a CSV File"
author: "Jose Zuniga"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Instructions

In this project, you're given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

* Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents. 

For the first player, the information would be: 

* Gary Hua, ON, 6.0, 1794, 1605 (1605 was calculated by using the pre-tournament opponents' ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.)

If you have questions about the meaning of the data or the results, please post them on the discussion forum. Data science, like chess, is a game of back and forth. 

*The chess rating system (invented by a Minnesota statistician named Arpad Elo) has been used in many other contexts, including assessing relative strength of employment candidates by human resource departments.*

### (1) Import Data into R

```{r}
tournamentinfo <- read.csv(paste0("https://raw.githubusercontent.com/jzuniga123",
                                  "/SPS/master/DATA%20607/tournamentinfo.txt"), header=F)
head(tournamentinfo)
```
The [raw data](https://github.com/jzuniga123/SPS/blob/master/DATA%20607/tournamentinfo.txt) has hyphens across every $(3n+1)th$ row to serparate the data for each subject which is written across both the $(3n+2)th$ row and the $(3n+3)th$ row.


### (2) Remove First Few Rows which Contain Headings.
```{r}
tournamentinfo2 <- tournamentinfo[-c(1:3),]
```

### (3) Pull Out Every $(3n+2)th$ and $(3n+3)th$ Row.
```{r}
n <- length(tournamentinfo2)
row1 <- tournamentinfo2[seq(2, n, 3)]
row2 <- tournamentinfo2[seq(3, n, 3)]
```

### (4) Extract Specified Data with Regular Expressions.
```{r}
library(stringr)
P_Number <- as.integer(str_extract(row1, "\\d+"))
P_Name <- str_trim(str_extract(row1, "(\\w+\\s){2,3}"))
P_State <- str_extract(row2, "\\w+")
P_Points <- as.numeric(str_extract(row1, "\\d+\\.\\d+"))
P_PreRating <- as.integer(str_extract(str_extract(row2, "[^\\d]\\d{3,4}[^\\d]"), "\\d+"))
Opponents <- str_extract_all(str_extract_all(row1, "\\d+\\|"), "\\d+")
```

### (5) Calculate Average of Opponents' Scores.
```{r}
O_PreRating <- numeric(n / 3)

for (i in 1:(n / 3)) { 
  O_PreRating[i] <- mean(P_PreRating[as.numeric(unlist(Opponents[P_Number[i]]))]) 
}
```

### (6) Create a Data Frame
```{r}
csv <- data.frame(P_Name, P_State, P_Points, P_PreRating, O_PreRating); csv
```

### (7) Export Data to CSV File
```{r eval=FALSE}
write.table(csv, file = "DATA_607_Project1.csv", sep = ",", col.names = T, row.names = F)
```
A copy of the exported .csv file can be found [here](https://github.com/jzuniga123/SPS/blob/master/DATA%20607/DATA_607_Project1.csv).

## Data Dictionary

* Pair Num for our purposes is just an ID number.  In "tournament scheduling" software, it would be used to communicate next round pairings -- "1" plays White vs. "8", etc.
* RTG (Pre->Post) is the rating before the event, and the rating after the event, with the delta based on performance (based on score and strength of opponents).  This is based on a formula by a Minnesota statistician named Arpad Elo, and has been used to rank professional tennis players, and human resource job candidates.  You should use the pre-rating (1794 for Gary Hua) in the assignment. Gary gained 23 points, a good result.
* ON means Ontario (State or province code).
* N:2 indicates the player's current highest norm attained (e.g. master, expert, class A, like martial arts belts: Black, Brown, etc.)
* W/B indicates whether player was assigned to play White (moves first) or Black in that game.