---
title: "NYC Mortality Statistics"
author: "Jose Zuniga"
output:
  html_document:
    toc: yes
    toc_depth: '4'
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Final Project Instructions

Describe your motivation for performing this analysis (2). Have a recognizable data science workflow (4.a - 4.e, 5.a - 5.c). Include data from at least two different types of data sources (4.b.i, 4.e.iv.3, 5.a.i). Include at least one data transformation operation (4.b.ii, 5.a.ii). Include at least one statistical analysis and at least one graphic that describes or validates the data (4.d, 5.b). Includes at least one graphic that supports the conclusion (4.e, 5.b). Includes at least one statistical analysis that supports the conclusion (4.e, 5.c). Includes at least one feature not covered in class (5.a - 5.b). Deliver the presentation in the allotted time of 3 to 5 minutes. Show at least one challenge encountered in code or data, and what was done when the challenge was encountered (4.e.iv). Ensure the audience comes away with a clear understanding of the motivation for undertaking the project. Ensure the audience comes away with a clear understanding of at least one insight gained, the conclusion you reached, or the hypothesis you confirmed. Deliver the submit self-contained code and data. Deliver fully reproducible results. Ensure all of the delivered code runs without errors. Deliver code and conclusions using a reproducible research tool. Deliver draft project proposal, project, and presentation on time. 

***

# 2. Purpose of Project

In April of 2015 there was a murder in New York City. There was a surge of police activity. Then, when crime scene investigation was done, the body remained covered by a white sheet and guarded by a police officer for a few hours until it was picked up by an ambulance. The strange thing is, a few news outlets reported that "EMS rushed him to St. Barnabas Hospital, where he was pronounced dead, cops said." The whole situation was disturbing, but these new reports also made me wonder how crime statistics are reported. Was this death reported in the statistics for the precinct of the hospital (48th) or the precinct of the crime (46th)? I never did the digging necessary to find out.

Then, in September of 2016, I read something about a similar situation. There was a story about the time of death reported by the NYC Police Department. The report said "Police officials initially said that Graham died at Montefiore Medical Center. His death certificate says he was shot at 3:01 p.m. and died at 3:53 p.m. According to his family, the video indicates he was already dead, and contradicts the city account." Although this story was also disturbing in-and-of itself, it also made me wonder if the death was reported in the statistics for the precinct of the hospital (52nd) or the precinct of the crime (47th)? I also wondered if the total number of mortalities reported in NYC police statistics match the total number of mortalities reported in NYC Hospitals statistics.

# 3. Question(s)

Do NYC Police Precincts and NYC Hospitals report different mortality figures? This question will be examined in general terms (whether NYC Police Precincts and NYC Hospitals report the same total mortality figures in their respective statistics) and much more specific terms (whether mortality statistics are skewed toward police precincts near hospitals).

```{r  message=F, warning=F}
library(XML)
library(httr)
library(tidyr)
library(dplyr)
library(RCurl)
library(ggmap)
library(ggplot2)
library(stringr)
library(jsonlite)
```

# 4.	General Question

Do NYC Police Precincts and NYC Hospitals report the same total mortality figures in their statistics?

## a. Obtain Data

### i.  NYC Police Department (PD)

  + [Seven Major Felony Offenses](http://www.nyc.gov/html/nypd/html/analysis_and_planning/historical_nyc_crime_data.shtml). Statistical breakdown by citywide, borough, and precinct 2000 - 2015.

### ii.  NYC Department of Health and Mental Hygiene (DOHMH)

  + [New York City Leading Causes of Death](https://data.cityofnewyork.us/api/views/jb7j-dtam/rows.csv?accessType=DOWNLOAD). The leading causes of death by sex and ethnicity in New York City 2007 - 2011.

## b.	Scrub Data

### i. Load Data Sets

```{r}
Mortality_1 <- read.csv(paste0("https://raw.githubusercontent.com/jzuniga123/SPS/master/",
                               "DATA%20607/DATA_607_Final_Project_1.csv"), stringsAsFactors = F)

Mortality_2 <- read.csv(paste0("https://raw.githubusercontent.com/jzuniga123/SPS/master/",
                               "DATA%20607/DATA_607_Final_Project_2.csv"), stringsAsFactors = F)
```

### ii. Clean, Preprocess, and Reduce Data Sets

```{r}
Precincts_1 <- Mortality_1 %>%
  filter(grepl('MURDER', CRIME)) %>%
  gather(xYear, Count, 3:18) %>%
  na.omit(Count) %>%  
  mutate(Year = as.integer(sub("X", "", xYear))) %>%
  group_by(Year) %>%
  summarise(Total = sum(Count))

Hospital_1 <- Mortality_2 %>% 
  filter(grepl('HOMICIDE', Cause.of.Death)) %>% 
  group_by(Year) %>% 
  summarise(Total = sum(Count))
```

## c. Explore Data

### i. Determine the Task

Test the significance of the difference between pairs. 

## d. Model Data

### i.  Choose the Technique

```{r}
shapiro.test(unlist(Hospital_1))
shapiro.test(unlist(Precincts_1))
bartlett.test(Hospital_1,Precincts_1)
```
Given fairly normal and homoscedastic matched pairs, and assuming independent cases (years), the **paired t-test** would be appropriate. The paired t-test compares two populations of paired data that are correlated due to said pairing. It uses the mean and standard deviation of the differences to calculate $t=\frac { \bar { d }  }{ s(d)/{ \sqrt { n }  } }$. The null hypothesis of the test is that the true mean difference is zero.

### ii.  Use Algorithm to Perform Analysis

```{r}
(Compare_1 <- full_join(Precincts_1, Hospital_1, by = "Year", suffix = c(".NYPD", ".DOHMH")))
visualize <- rbind(cbind(Hospital_1, Source = "DOHMH"), cbind(Precincts_1, Source = "NYPD"))
ggplot() + geom_area(aes(x = Year, y = Total, fill = Source), data = visualize, 
           stat="identity", position = position_dodge(width = 0), alpha= I(0.5))
Compare_2 <- right_join(Precincts_1, Hospital_1, by = "Year", suffix = c(".NYPD", ".DOHMH"))
data.frame(Compare_2, Difference = matrix((unlist(Compare_2[3] - Compare_2[2])), ncol=1),
           Factor = matrix((unlist(Compare_2[3] / Compare_2[2])), ncol=1))
t.test(Compare_2[3] - Compare_2[2])
```

## e.  Interpret Data

### i. Comparison

The data reported by the NYC PD and for the NYC DOHMH is conflicting and flawed. The NYC PD data shows `r nrow(Precincts_1)` years of data with an average of `r round(mean(unlist(Precincts_1[2])), 0)` while the NYC DOHMH data shows `r nrow(Hospital_1)` years of data with an average of `r round(mean(unlist(Hospital_1[2])), 0)`. When looking at the years where both have data available, the average difference is `r round(mean(unlist(Compare_2[3] - Compare_2[2])), 0)` which is very strange. The NYC DOHMH data is inflated by a constant factor of about `r round(mean(unlist(Compare_2[3] / Compare_2[2])), 2)`.

### ii.  Statistical Test

There is something wrong here that does not justify even discussing the **t-test** results. 

### iii. Findings

Occam's razor demands that the fewest assumptions be made in the interpretation of these findings. Therefore, it is likely that something is wrong in the data. Further investigation is necessary.

### iv.	Investigate Further

#### 1. NYC PD Freedom of Information Law Request

> OIGNYPDFoil@oignypd.nyc.gov

> FOIL Request #: 2016.0056

> This is in response to your October 27, 2016 (received by the Office of the Inspector General for the NYPD ["OIG-NYPD"] on October 27, 2016) request, pursuant to the Freedom of Information Law, for "formal definition of the 'MURDER & NON NEGL. MANSLAUGHTER' CompStat category."
 
> The New York Police Department uses the New York State Penal Law to define all crimes/offenses/infractions referred to in their CompStat 2.0 reporting data. 
 
> Please note that the OIG-NYPD is not part of the NYPD.  For records kept by the NYPD, you may make an electronic FOIL request to the NYPD's Records Access Officer at http://www.nyc.gov/html/nypd/html/legal_matters/dclm_doc_production_foil.shtml.
 
> This letter represents our complete response to your request. Should you wish to appeal this determination, you must send written notice within thirty days to Asim Rehman, General Counsel at the Office of the Inspector General for the New York City Police Department, City of New York Department of Investigation, 80 Maiden Lane, 14th Floor, New York, New York 10038.

**New York State Penal Law**

S 125.00 Homicide defined: Homicide means conduct which causes the death of a person or an unborn child with which a female has been pregnant for more than twenty-four weeks under circumstances constituting murder, manslaughter in the first degree, manslaughter in the second degree, criminally negligent homicide, abortion in the first degree or self-abortion in the first degree.

Section Offense	                                      Class
------  --------------------------------------------  -----------
125.10	Criminally negligent homicide.	              E FELONY
125.11	Aggravated criminally negligent homicide.	    C FELONY
125.12	Vehicular manslaughter in the second degree.  D FELONY
125.13	Vehicular manslaughter in the first degree.   C FELONY
125.14	Aggravated vehicular homicide.	              B FELONY
125.15	Manslaughter in the second degree.	          C FELONY
125.20	Manslaughter in the first degree.	            B FELONY
125.21	Aggravated manslaughter in the second degree. C FELONY
125.22	Aggravated manslaughter in the first degree.  B FELONY
125.25	Murder in the second degree.	                A-I FELONY
125.26	Aggravated murder.	                          A-I FELONY
125.27	Murder in the first degree.	                  A-I FELONY
125.40	Abortion in the second degree.	              E FELONY
125.45	Abortion in the first degree.	                D FELONY
125.50	Self-abortion in the second degree.	          B MISD
125.55	Self-abortion in the first degree.	          A MISD
125.60	Issuing abortional articles.	                B MISD
-----------------------------------------------------------------

#### 2. NYC DOH Freedom of Information Law Request

> recordsaccess@health.nyc.gov

> FOIL Request #216FR02936

> The NYC Department of Health and Mental Hygiene has received your Freedom of Information Law request and assigned it the control number noted above.  The data you seek for 2000 - 2014 is available online at the Department's EpiQuery website:  https://sasebiweb200.health.dohmh.nycnet/epiquery/VS/index.html .  If you select "Mortality - by select causes" and then click on the "SUBMIT" button, the data will appear.  Please note that 2015 data has not yet been finalized.

> The definition of "homicide" is also available online at http://www1.nyc.gov/assets/doh/downloads/pdf/ip/ip-homicides-in-new-york-city.pdf.
 
> This concludes the Department's response to your Freedom of Information Law request.

**International Classification of Disease (ICD)-10**

Category  Condition
--------  -----------------------------------------------------------
X85       Assault by drugs, medicaments and biological substances
X86       Assault by corrosive substance
X87       Assault by pesticides
X88       Assault by gases and vapors
X89       Assault by other specified chemicals and noxious substances
X90       Assault by unspecified chemical or noxious substance
X91       Assault by hanging, strangulation and suffocation
X92       Assault by drowning and submersion
X93       Assault by handgun discharge
X94       Assault by rifle, shotgun and larger firearm discharge
X95       Assault by other and unspecified firearm discharge
X96       Assault by explosive material
X97       Assault by smoke, fire and flames
X98       Assault by steam, hot vapors and hot objects
X99       Assault by sharp object
Y00       Assault by blunt object
Y01       Assault by pushing from high place
Y02       Assault by pushing or placing victim before moving object
Y03       Assault by crashing of motor vehicle
Y04       Assault by bodily force
Y05       Sexual assault by bodily force
Y06       Neglect and abandonment
Y07       Other maltreatment syndromes
Y08       Assault by other specified means
Y09       Assault by unspecified means
Y87.1     Sequela of assault ^i^
---------------------------------------------------------------------
^i^ The sequela include conditions reported as such, or occurring 
as "late effects" one year or more after the originating event.

#### 3. Scrub, Explore, Model

```{r}
Mortality_3 <- read.csv(paste0("https://raw.githubusercontent.com/jzuniga123/SPS/master/",
                               "DATA%20607/DATA_607_Final_Project_3.csv"), stringsAsFactors = F)

Hospital_2 <- Mortality_3[1:15, 2:3]
Hospital_2[1] <- as.integer(unlist(Hospital_2[1]))
colnames(Hospital_2) <- colnames(Hospital_1)
shapiro.test(unlist(Hospital_2))
bartlett.test(Hospital_2, Precincts_1)
visualize <- rbind(cbind(Hospital_2, Source = "DOHMH"), cbind(Precincts_1, Source = "NYPD"))
ggplot() + geom_area(aes(x = Year, y = Total, fill = Source), data = visualize, 
           stat="identity", position = position_dodge(width = 0), alpha= I(0.5))
Compare_3 <- right_join(Precincts_1, Hospital_2, by = "Year", suffix = c(".NYPD", ".DOHMH"))
data.frame(Compare_3, Difference = matrix((unlist(Compare_3[3] - Compare_3[2])), ncol=1),
           Factor = matrix((unlist(Compare_3[3] / Compare_3[2])), ncol=1))
t.test(Compare_3[3] - Compare_3[2])
```

#### 4. Comparison

These data are much better. The new NYC DOHMH data shows `r nrow(Hospital_2)` years of data with an average of `r round(mean(unlist(Hospital_2[2])), 0)`. When looking at the years where both have data available, the average difference is `r round(mean(unlist(Compare_3[3] - Compare_3[2])), 0)`. The new NYC DOHMH data is still inflated, but now only by a factor of `r round(mean(unlist(Compare_3[3] / Compare_3[2])), 2)`.

#### 5. Statistical Test

For the t-test, the null hypothesis is set to a zero difference, $d = 0$, between the NYC PD murder statistics and NYC DOHMH mortality statistics (this would indicate variation is due to reporting discrepancies), and the alternate hypothesis to an actual difference, $d \neq 0$, existing between the two agencies. Failure to reject the null hypothesis will be assessed on a $p$-value of $0.05$ such that $H_{ 0 }: p > 0.05; H_{ A }: p \le 0.05$. The test results indicate the existence of a $p$-value $=`r format(t.test(Compare_3[3] - Compare_3[2])$p.value, scientific=F)` < 0.05$ indicating that the probability of observing a sample $t$ statistic with `r t.test(Compare_3[3] - Compare_3[2])$parameter` degrees of freedom as extreme as the test statistic `r t.test(Compare_3[3] - Compare_3[2])$statistic` is extremely low. Therefore, the null hypothesis is rejected. 

### v.  Findings

The data supports the conclusion that an actual difference exists between NYC PD mortality statistics and NYC DOHMH mortality statistics. This begs the question about where this statistically significant and important difference in mortalities is coming from. Examining the data definitions, it seems reasonable to check if a sequela lag issue (mortality reported as such, or occurring as "late effects" one year or more after the originating event) can be the reason.

### vi.	Refine Further

```{r}
(sequela <- data.frame("Year" = Compare_3[1] - 1, Compare_3[3] - Compare_3[2]))
sequela_lag <- Precincts_1[1:14, ]
sequela_lag[ , 2] <-  sequela_lag[ , 2] + sequela[2:15 ,2]
visualize <- rbind(cbind(Hospital_2[1:14,], Source = "DOHMH"), cbind(sequela_lag, Source = "NYPD"))
ggplot() + geom_area(aes(x = Year, y = Total, fill = Source), data = visualize, 
           stat="identity", position = position_dodge(width = 0), alpha= I(0.5))
Compare_4 <- left_join(sequela_lag, Hospital_2, by = "Year", suffix = c(".NYPD", ".DOHMH"))
data.frame(Compare_4, Difference = matrix((unlist(Compare_4[3] - Compare_4[2])), ncol=1),
           Factor = matrix((unlist(Compare_4[3] / Compare_4[2])), ncol=1))
t.test(Compare_4[3] - Compare_4[2])
```

#### 1. Comparison

Lagging the difference between agencies one year (although it can be greater), the modified NYC PD data shows `r nrow(sequela_lag)` years of data with an average of `r round(mean(unlist(sequela_lag[2])), 0)`. When looking at the years where both have data available, the average difference is `r round(mean(unlist(Compare_4[3] - Compare_4[2])), 0)`. The new modified data is still showing some variation, but now the mean factor is `r round(mean(unlist(Compare_4[3] / Compare_4[2])), 2)`. The differences basically cancel out.

#### 2. Statistical Test

For the t-test, the null hypothesis is set to a zero difference, $d = 0$, between the NYC PD mortality statistics **including sequela** and NYC DOHMH mortality statistics (this would indicate variation is due to reporting discrepancies), and the alternate hypothesis to an actual difference, $d \neq 0$, existing between the two agencies. Failure to reject the null hypothesis will be assessed on a $p$-value of $0.05$ such that $H_{ 0 }: p > 0.05; H_{ A }: p \le 0.05$. The test results indicate the existence of a $p$-value $=`r format(t.test(Compare_4[3] - Compare_4[2])$p.value, scientific=F)` > 0.05$ indicating that the probability of observing a sample $t$ statistic with `r t.test(Compare_4[3] - Compare_4[2])$parameter` degrees of freedom as extreme as the test statistic `r t.test(Compare_4[3] - Compare_4[2])$statistic` is extremely high. Therefore, we fail to reject the null hypothesis. 

### vii. Findings

The data supports the conclusion that no difference exists between NYC PD mortality statistics and NYC DOHMH mortality statistics, or in other words, the variation in difference is due to reporting differences. Now the question about where mortalities are reported--at the precinct of the hospital or the precinct of the crime--- will be addressed.

# 5.	Specific Question

Are mortality statistics skewed toward police precincts near hospitals?

## a. Obtain, Scrub, Explore Data

### i.  Google Geocoding API

The `geocode()` function from the `ggmap` library is extremely efficient but only returns longitude and latitude coordinates. Using *Google's* geocoding API returns much more information. The API allows 2,500 free requests per day and returns results in `XML` or `JSON` format. The API has a limit of 50 requests per second.
```{r cache=T}
NYC_Precincts <- read.csv(paste0("https://raw.githubusercontent.com/jzuniga123/SPS/master/",
                               "DATA%20607/DATA_607_Final_Project_4.csv"), stringsAsFactors = F)

NYC_Hospitals <- read.csv(paste0("https://raw.githubusercontent.com/jzuniga123/SPS/master/",
                               "DATA%20607/DATA_607_Final_Project_5.csv"), stringsAsFactors = F)

Geocode_1 <- matrix(NA, nrow(NYC_Precincts), 13)
for (i in 1:nrow(NYC_Precincts)) {
  Geocode_1[i, 1] = NYC_Precincts[i, 1]
  url <- "https://maps.googleapis.com/maps/api/geocode/json?address="
  url <- str_replace_all(paste0(url, NYC_Precincts[i, 3],", ", NYC_Precincts[i, 4], ", NY"), "\\s", "+")
  json <- GET(url)
  json <- rawToChar(json$content)
  json <- fromJSON(json)
  geocode <- data.frame(json)
  Geocode_1[i, c(2:3)] <- if(!is.null(geocode[[1,3]]$lat[1])) {
    c(geocode[[1,3]]$lat[1], geocode[[1,3]]$lng[1])
  } else if (length(unlist(geocode[[1,3]][2])) == 1){
    c(unlist(geocode[[1,3]][1]), unlist(geocode[[1,3]][2]))
  } else { unlist(geocode[[1,3]][1]) }
  Geocode_1[i, 4] <- geocode[[2]][1]
  Geocode_1[i, c(5:13)]  <- data.frame(geocode[[1]][1])[1:9, 1]
  while (nchar(Geocode_1[i, 13]) != 5 | is.na(Geocode_1[i, 13])) {
    if (nchar(Geocode_1[i, 12]) != 5 | is.na(Geocode_1[i, 12])) {
      Geocode_1[i, c(10:13)] <- Geocode_1[i, c(9:12)]
      Geocode_1[i, 8] <- Geocode_1[i, 7]
    } else {
      Geocode_1[i, c(10:13)] <- Geocode_1[i, c(9:12)]
      Geocode_1[i, 9] <- Geocode_1[i, 8]
    }
  }
  Sys.sleep(1 / 3) # Make no more than 3 requests per second
}
head(Geocode_1[, 2:4])
Geocode_2 <- matrix(NA, nrow(NYC_Hospitals), 13)
for(i in 1:nrow(NYC_Hospitals)) {
  Geocode_2[i, 1] = NYC_Hospitals[i, 1]
  url <- "https://maps.googleapis.com/maps/api/geocode/json?address="
  url <- str_replace_all(paste0(url, NYC_Hospitals[i, 2], ", ",
          NYC_Hospitals[i, 3],", NY ", NYC_Hospitals[i, 5]), "\\s", "+")
  json <- GET(url)
  json <- rawToChar(json$content)
  json <- fromJSON(json)
  geocode <- data.frame(json)
  Geocode_2[i, c(2:3)] <- if(!is.null(geocode[[1,3]]$lat[1])) {
    c(geocode[[1,3]]$lat[1], geocode[[1,3]]$lng[1])
  } else if (length(unlist(geocode[[1,3]][2])) == 1){
    c(unlist(geocode[[1,3]][1]), unlist(geocode[[1,3]][2]))
  } else { unlist(geocode[[1,3]][1]) }
  Geocode_2[i, 4] <- geocode[[2]][1]
  Geocode_2[i, c(5:13)]  <- data.frame(geocode[[1]][1])[1:9, 1]
  while (nchar(Geocode_2[i, 13]) != 5 | is.na(Geocode_2[i, 13])) {
    if (nchar(Geocode_2[i, 12]) != 5 | is.na(Geocode_2[i, 12])) {
      Geocode_2[i, c(10:13)] <- Geocode_2[i, c(9:12)]
      Geocode_2[i, 8] <- Geocode_2[i, 7]
    } else {
      Geocode_2[i, c(10:13)] <- Geocode_2[i, c(9:12)]
      Geocode_2[i, 9] <- Geocode_2[i, 8]
    }
  }
  Sys.sleep(1 / 2) # Make no more than 3 requests per second
}
head(Geocode_2[, 2:4])
```

### ii. Calculate Distances

Given latitudes $\lambda_{ 1 }, \lambda_{ 2 }$ and longitudes $\varphi _{ 1 }, \varphi _{ 2 }$ in radians, as well as the Earth's mean radius in miles of $R=3959$, the distance between two points in miles can be calculated trigonometrically using the below formula.

$$d=\arccos { \left( \sin { \left( { \lambda  }_{ 1 } \right)  } \sin { \left( { \lambda  }_{ 2 } \right)  } +\cos { \left( { \lambda  }_{ 1 } \right)  } \cos { \left( { \lambda  }_{ 2 } \right)  } \cos { \left( \varphi _{ 1 }-\varphi _{ 2 } \right)  }  \right) R }$$

```{r}
Distance <- matrix(NA, nrow(Geocode_1), nrow(Geocode_2))
R = 3959 # Earth mean radius in miles
for (i in 1:nrow(Geocode_1)){
  lat1 = as.double(Geocode_1[i, 2]) * pi / 180
  lng1 = as.double(Geocode_1[i, 3]) * pi / 180
  for (j in 1:nrow(Geocode_2)) {
    lat2 = as.double(Geocode_2[j, 2]) * pi / 180
    lng2 = as.double(Geocode_2[j, 3]) * pi / 180
    Distance[i, j] = acos(sin(lat1) * sin(lat2) + cos(lat1) * cos(lat2) * cos(lng2 - lng1)) * R
  }
}
NYC_Precincts <- Distance %>% as.data.frame() %>% 
  mutate(Min_Dist = apply(.[, apply(., 1, is.numeric)], 1, min)) %>% 
  select(ncol(Distance) + 1) %>% 
  cbind(Geocode_1[ , 1:3], ., Geocode_1[ , c(4,9,13)])
colnames(NYC_Precincts) <- c("Precinct", "Lat", "Lng", "Min_Dist", "Address", "Borough", "ZIP")
NYC_Hospitals <- Distance %>% t() %>% as.data.frame() %>% 
  mutate(Min_Dist = apply(.[, apply(., 2, is.numeric)], 1, min)) %>% 
  select(nrow(Distance) + 1) %>% 
  cbind(Geocode_2[ , 1:3], ., Geocode_2[ , c(4,9,13)])
colnames(NYC_Hospitals) <- c("Precinct", "Lat", "Lng", "Min_Dist", "Address", "Borough", "ZIP")
Precincts_2 <- Mortality_1 %>%
  filter(grepl('MURDER', CRIME) & PCT != "DOC") %>%
  gather(Year, Count, 3:18) %>%
  mutate(Count = ifelse(is.na(Count), 0, Count)) %>% 
  mutate(Year = as.integer(sub("X", "", Year))) %>%
  rename(Precinct = PCT) %>% 
  group_by(Precinct) %>%
  summarise(Total = sum(Count))
```

### iii.	Determine the Task

Test the correlation between mortality and distance.

## b.	Model Data

### i.	Choose the Technique

Correlation coefficients measure the strength of association between two variables. The most common correlation coefficient, called the Pearson product-moment correlation coefficient, measures the strength of the linear association between variables.

### ii.	Use Algorithm to Perform Analysis

```{r warning=F, message=F}
ggmap(get_map(location=geocode("NYC"), source="google", maptype="roadmap")) + 
  geom_point(aes(y = as.numeric(levels(Lat))[Lat], x = as.numeric(levels(Lng))[Lng]), 
  data = NYC_Precincts, alpha = 0.5, color="#009999", size = 2) +
  geom_point(aes(y = as.numeric(levels(Lat))[Lat], x = as.numeric(levels(Lng))[Lng]), 
  data = NYC_Hospitals, alpha = 0.5, color="#FF6699", size = 2)
ggmap(get_map(location=geocode("Brooklyn, NY"), source="google", maptype="roadmap", zoom=11)) + 
  geom_point(aes(y = as.numeric(levels(Lat))[Lat], x = as.numeric(levels(Lng))[Lng]), 
  data = NYC_Precincts, alpha = 0.5, color="#009999", size = 3) +
  geom_point(aes(y = as.numeric(levels(Lat))[Lat], x = as.numeric(levels(Lng))[Lng]), 
  data = NYC_Hospitals, alpha = 0.5, color="#FF6699", size = 3)
Compare_5 <- left_join(NYC_Precincts, Precincts_2, by = "Precinct")
cor(Compare_5[, c(4,8)])[[2]]
ggplot(Compare_5, aes(Min_Dist, Total)) + 
  geom_point(aes(colour = factor(Borough))) + 
  labs(x = "Miles to Nearest Hospital", y = "Total Mortality") +
  labs(colour = "Borough", title = "Mortality Relative to Proximitry") +
  theme(plot.title = element_text(hjust = 0.5))
(Mortality_4 <- Compare_5 %>% 
  group_by(Borough) %>% summarise(Total = sum(Total)) %>% 
  cbind(., Population = c(1385107, 2504710, 1585874, 2230541, 468730)) %>% 
  mutate(Frequency = paste0(round(100 * Total / Population, 2),"%")))
```

## c.	Interpret Data

### i.	Comparison

The police precincts and hospitals plotted on a map show relatively even spacing between police precincts and hospitals. Zooming in further also shows how close police precincts and hospitals are to each other. The scatter plot makes this relationship even more obvious with the vast majority of distances clustered in the area representing less than two miles. It is also worth noting the difference between mortality totals and mortality frequency relative to population in each borough, as perception could be biased by looking solely at totals.

### ii.	Statistical Test

The correlation between mortality and distance to nearest hospital is negative with $r=`r cor(Compare_5[, c(4,8)])[[2]]`$. This is an extremely weak correlation which does not support the conclusion that there exists a relationship between mortality figures reported from precincts closer to hospitals.

### iii.	Findings

The data supports the conclusion that no relationship exists between mortality figures reported from precincts closer to hospitals and those reported from precincts further from hospitals.

# 6.  Conclusion

In general, NYC police statistics appear to match NYC hospital statistics when it comes to mortalities. Differences in data reported by each agency appear to be attributable to how sequela is handled. What this analysis also shows is both the disparity between the quality of data and the importance of working through initial findings critically.

Specifically, there is no evidence to support the hypothesis that mortality figures are skewed toward precincts closer to hospitals. Therefore, it is highly unlikely that mortality is reported at the precinct of the hospital rather than the precinct of the crime. It is worth noting both that the number of crimes, police precincts, and hospitals tend to higher in high density areas and that hospitals do not necessarily lie in the precinct to which they are closest. Yet given the questions raised and the findings, a micro level examination of those factors is unwarranted in this analysis.

# 7.  Comment

The events that inspired this analysis were tragedies. Trying to explain what happened is not easy, if at all possible. At least now some possibilities can be ruled out. Yet regardless of what transpired and why, I would like to add this note extending condolences and respect to the families of the individuals involved. 

# 8.  References

http://nypost.com/2015/04/06/daily-blotter-485/

http://www.nydailynews.com/new-york/video-shows-ramarley-graham-carried-police-shooting-article-1.2810982

http://www.ats.ucla.edu/stat/mult_pkg/whatstat/default.htm

http://www.statstutor.ac.uk/resources/uploaded/paired-t-test.pdf

http://www.statisticssolutions.com/manova-analysis-paired-sample-t-test/

http://stattrek.com/hypothesis-test/paired-means.aspx?Tutorial=AP

http://stattrek.com/statistics/correlation.aspx

http://www.nyc.gov/html/oignypd/pages/about/foil-request.shtml

http://ypdcrime.com/penal.law/article125.htm#p125.27

http://www.nycourts.gov/judges/cji/2-PenalLaw/125/art125hp.shtml

https://www1.nyc.gov/site/doh/about/ogc-foil.page

https://a816-healthpsi.nyc.gov/epiquery/VS/index.html

http://apps.who.int/classifications/apps/icd/icd10online2005/fr-icd.htm?gx85.htm+

http://apps.who.int/classifications/apps/icd/icd10online2005/fr-icd.htm?gy85.htm+

https://en.wikipedia.org/wiki/ICD-10_Chapter_XX:_External_causes_of_morbidity_and_mortality

https://a816-healthpsi.nyc.gov/epiquery/CHS/uhf-zip-information.pdf

https://www1.nyc.gov/assets/doh/downloads/pdf/ah/zipcodetable.pdf
 
http://www.health.ny.gov/regulations/hcra/provider/provhosp.htm

https://en.wikipedia.org/wiki/Acute_care
 
http://www.nyc.gov/html/nypd/html/home/precincts.shtml
 
https://developers.google.com/maps/documentation/geocoding

http://www.sunearthtools.com/tools/distance.php

https://www.citypopulation.de/php/usa-newyorkcity.php